Space astrometry of the very massive $\sim$$150\,M_\odot$ candidate runaway star VFTS682

### 2018-09-28 MNRASformat/referee_report_answer.txt

=========================================================================================

1. Sec. 1,  paragraph 4: in the sentence beginning with "N-body simulations indicate that…”, it would be clearer to add the phrase “through dynamical ejections” after the word “expected”.

ANSWER: =========================================================================================

We agree that this would be more clear, and have modified the sentence accordingly. The text now reads: "N- body simulations indicate that the *dynamical* ejection of..."

@@

=================================================================================================

2. It is not immediately clear to me how the widths of the green vertical bands in Fig. 1 are estimated. It is important to explain this in the text - these bands are important to demonstrate the likeliness of VFTS682 for being an early dynamical runaway from R136. Perhaps similar bands can also be indicated for VFTS16 and VFTS72 in the panels of Fig. 1, for comparing their cases with that of VFTS682.

ANSWER: =========================================================================================

Indeed, we agree that we should have stated this more clearly. We have added a clarification in the main text of Sec. 3 and a pointer to it in the figure caption.

@@

To answer the referee's question here: the green vertical band in the top panel is determined by the error bars on the apparent age of the star: the expected proper motion is the projected distance between the cluster and the star (~29pc) divided by the apparent age of the star (1+/-0.2Myr) which gives the v2D reported in Tab. 2. For the bottom panel we have assumed a width of 45 degrees centered in theta=0.

We have chosen not to include similar bands for VFTS16 and 72 since they would overlap with the one for VFTS682 in the bottom panel and we think this would make the figure less clear. We note that this information is also contained in Fig. 2 (in the prolongation backward in time of the observed proper motions and the proper motion error cone).

@@

6. Sec. 4, paragraph 4: "In their model, they relied on (dynamically driven) stellar mergers to explain the high masses of VFTS682 and the massive members of R136.” Note that the dynamical merger scenario is not instrumental in explaining members (either residing in or dynamically ejected from R136) up to initial stellar masses of 150 M_sun, due to the adopted IMF in the models of Banerjee et al. (2012) that extends up to 150 M_sun. It is necessary to explain any single star (in-cluster or ejected) beyond 150 M_sun, given these models: see Banerjee S. et al., 2012, MNRAS, 426, 1416 and Crowther P. A. et al., 2010, MNRAS, 408, 731. (In case > 150 M_sun stars would also derive directly from massive-star-formation process, stellar mergers would not be instrumental also for such “super-canonical” stars - for such stars, the dynamical ejection fraction approaches unity as in Banerjee, S. et al., 2012, ApJ, 746, 15 and Fujii, M. S., Portegies Zwart S., 2011, Science, 334, 1380.) Therefore, I suggest rephrasing the quoted sentence accordingly.

ANSWER: =========================================================================================

We agree with the referee on this point, and have modified the text to make the dynamical mergers less upfront.

We would prefer to still mention the mergers, to implicitly pose the question of whether these VMS are mergers or not (and we agree with the referee that this remains to be seen for VFTS682). While mergers are not instrumental in explaining their results, the fact that using an IMF up to 150Msun only already gave Banerjee et al. enough VMS stars both in and around the cluster suggests that if they had allowed for the formation of even more massive stars, then the efficiency of dynamically driven mergers together with stars initially more massivre than 150Msun might possibly result in an over-prediction of VMS in the region. We think that this is an interesting point were observations can put constraints on the physics of cluster dynamics and stellar mergers.

@@

=================================================================================================

7. Sec. 4, second to last paragraph: I recommend also citing the recent work Belczynski, K. et al., 2016, A&A, 594, A97 in the context of the (im)possibility of avoiding pair-instability SN, as invoked there. According to the above work, pair-instability SN can easily be avoided at the LMC metallicity for at least up to 150 M_sun ZAMS (the initially approx. 150 M_sun VFTS682 can easily be a non-merger product; see point 6).

ANSWER: =========================================================================================

We agree with the referee that the metallicity dependence of the pair-instability occurrence is also addressed in Belczynski et al. 2016, although it is a population synthesis study where the pulsational pair instability is prescribed rather than calculated, and therefore the reference suggested is not a primary one. The work of Belczynski et al. is based on the detailed simulations of Woosley 2017, http://adsabs.harvard.edu/abs/2017ApJ...836..244W, and we chose to add a citation to the primary source instead.

### 2018-09-28 MNRASformat/VFTS682_MNRASL.tex
added affiliation for Danny and cut fig. caption

\author[Renzo et al.]{M.~Renzo$^{1}$% \thanks{Corresponding author:
    % M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
  , S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, I.~Platais$^{3}$,
  R.~P.~van~der~Marel$^{3,4}$,
  \newauthor{E.~Laplace$^{1}$, J.~M.~Bestenlehner$^{5}$, C.~J.~Evans$^{6}$,
    V.~H\'enault-Brunet$^{7}$,  S.~Justham$^{8,9,1}$,
  }
\newauthor{A.~de~Koter$^{1}$,
      N.~Langer$^{10}$, F. Najarro$^{11}$, H.~Sana$^{12}$, F.~R.~N.~Schneider$^{13}$, J.~S.~Vink$^{14}$}}

@@

\begin{keywords}
  stars: astrometry, kinematics and dynamics, individual: VFTS682
\end{keywords}

\section{Introduction}
\label{sec:intro}

@@

 \hline     
      \multicolumn{5}{l}{\emph{HST proper motion of VFTS682 relative
      to R136}} \\
            \hline
      &$\delta\mu_\mathrm{RA, HST}$  &[$\masyr$] & $0.02\pm0.10$ & (5) \\
      &$\delta\mu_\mathrm{DEC, HST}$  &[$\masyr$] & $0.19\pm0.09$ &  (5) \\
       &$\delta\mu_\mathrm{HST}$  &[$\masyr$] & $0.19\pm0.09$ &  (5) \\
                  &$v_\mathrm{2D, HST}$  &[$\kms$] & $45\pm21$ & (5)\\  
                      &$\theta_\mathrm{HST}$  &[degrees] &   $-30_{-51}^{+24}$   & (1,6)\\  
                  %        &$v_\mathrm{2D, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
      \hline
      \multicolumn{5}{l}{\emph{Weighted average relative proper motion
      for VFTS682}}\\
      \hline
      &$\delta\mu_\mathrm{RA, avg}$  &[$\masyr$] & $0.08\pm0.07$ & (6) \\
      &$\delta\mu_\mathrm{DEC, avg}$  &[$\masyr$] & $0.14\pm0.07$ &  (6) \\
       &$\delta\mu_\mathrm{avg}$  &[$\masyr$] & $0.16\pm0.07$ &  (6) \\
      &$v_\mathrm{2D, avg}$  &[$\kms$] & $38\pm17$ & (6)\\
      % &$\theta_\mathrm{avg}$  &[degrees] &  \todo{TBD} $_{-}^{+}$   & (5)\\  
      \hline
       \multicolumn{5}{l}{\emph{Expected proper motion if ejected from
      R136 at age zero}} \\
      \hline
      &$v_\mathrm{2D}$  &[$\kms$] & $29\pm 6$ & (3) \\  
      &$\theta$  &[degrees] &  $\sim0$  & \\ 
       \hline
    \end{tabular}
    {\tiny The error on the RA and DEC positions, are of order
      $\sim$$10^{-2}\masyr$ in \emph{Gaia}
      DR2. Assuming a distance of 50\,kpc, the conversion from proper motion
    to projected velocities is such that $1\masyr$ corresponds to $237\kms$.
$\rho(\mu_\mathrm{RA},\mu_\mathrm{DEC})$ is the
      correlation coefficient. The position angle
      $\theta$ is defined such that $\theta=0$ for radial motion
      away from R136. We neglect the error bars on
      $\langle\mu_\mathrm{RA}\rangle_\mathrm{R136}$ and $\langle\mu_\mathrm{DEC}\rangle_\mathrm{R136}$ to determine the uncertainty
      on $\theta_{Gaia}$.
      (1)~\cite{brown:18},
      (2)~\cite{henault-brunet:12},
      (3)~\cite{bestenlehner:11},
      (4)~\cite{lennon:18}, 
      (5)~\cite{platais:18} and
      (6)~{\color{blue}this study}.
    }
  \end{center}
  \label{tab:vfts682}
\end{table}

@@

\begin{figure}%[tbp]
  \centering
  \includegraphics[width=0.35\textwidth]{figures/dist_mu_region.pdf}\\
  \vspace*{-5pt}
  \includegraphics[width=0.35\textwidth]{figures/angle}
  % \caption{\emph{Top panel}: distribution of OB-type and Wolf-Rayet stars in proper
  %   motion relative to R136 from \emph{Gaia} DR2. VFTS682 is not an outlier, but
  %   its relative proper motion matches the expected value (green
  %   vertical band) if it were indeed
  %   ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. %The top axis shows the conversion to $\kms$
  %   % assuming a distance of 50\,kpc.
  %   \emph{Bottom panel}:  distribution of
  %   angles between the relative proper motion direction and the radial
  %   direction from the center of the cluster to the star. The error bars from \emph{Gaia} for VFTS682 are large, but
  %   the best value is in agreement with the hypothesis of dynamical
  %   ejection (green vertical band). The peak at $\theta\simeq\pi/2$ is due to stars
  %   belonging to NGC2060. In both
  %   panels, the dark blue histograms contain 317 
  %   stars with error bars smaller than $0.1\,\mathrm{mas \
  %     yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
  %   lighter blue histograms contain 36 stars with error bars smaller than $0.05\,\mathrm{mas \
  %   yr^{-1}}$. }
  \caption{Distribution of OB-type and Wolf-Rayet stars in proper
    motion relative to R136 (top panel) and proper motion position angle
    (bottom panel), from \emph{Gaia} DR2. Although VFTS682 is not an outlier, its relative
    proper motion matches the value expected for an early dynamical
    ejection \newtext{(see \Secref{sec:results})}. In both
    panels, the dark blue histograms contain 317 
    stars with error bars smaller than $0.1\,\mathrm{mas \
      yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc and the
    lighter blue histograms contain 36 stars with error bars smaller than $0.05\,\mathrm{mas \
    yr^{-1}}$. The peak at $\theta\simeq\pi/2$ in the bottom panel is due to stars
  belonging to NGC2060. }
  \label{fig:dist}
\end{figure}

@@

% We list the relative proper motion in \Tabref{tab:vfts682}. 
\vspace*{-30pt}
\section{The kinematics of VFTS682}
\label{sec:results}

@@

If confirmed, it means that isolated star formation is
\emph{not} required to explain the isolation of VFTS682. Its proper motion suggests that it was ejected from the cluster R136
$0.7\pm0.3$\,Myr ago, which is compatible with the evolutionary age of
the star. If the
cluster age \citep[$\lesssim2$\,Myr,][]{crowther:10, sabbi:12} is
indeed smaller than the shortest stellar lifetime
\citep[$\sim$3\,Myr,][]{brott:11,kohler:15, zapartas:17}, the ejection of VFTS682
from the disruption of a \newtext{massive} binary by a supernova is excluded. %  since the region is too young for stars
% to have experienced core-collapse already. 
The kinematic age we infer is smaller than the kinematic age of
$\sim$1.5\,Myr for VFTS16 found in \cite{lennon:18}, which indicates
that VFTS682 was ejected later than VFTS16, and potentially later than
VFTS72 too.

@@

\vspace*{-20pt}
\section*{Affiliations}
 \tiny
\noindent $^{1}${Astronomical Institute Anton Pannekoek, University of
    Amsterdam, 1098 XH Amsterdam, The Netherlands} 
  $^{2}$ {ESA, European Space Astronomy Centre, Apdo. de Correos 78,
    E-28691 Villanueva de la Ca\~nada, Madrid, Spain} 
 $^{3}$ {
 %Center for Astrophysical Sciences, 
 Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}
  $^{4}$ {Space Telescope Science Institute, 3700 San Martin Drive,
    Baltimore, MD 21218, USA}
  $^{5}${Department of Physics and Astronomy, Hicks Building,
    Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}
  $^{6}${UK Astronomy Technology Centre, Royal Observatory Edinburgh, Blackford Hill, Edinburgh, EH9 3HJ, UK}
  $^{7}$ {National Research Council, Herzberg Astronomy \&
    Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7,
    Canada}
  $^{8}$ {School of Astronomy \& Space Science, University of the Chinese
    Academy of Sciences, Beijing 100012, China}
  $^{9}$ {National Astronomical Observatories, Chinese Academy of
    Sciences, Beijing 100012, China}
  $^{10}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn,
    Auf dem H\"ugel 71, 53121, Bonn, Germany}
  $^{11}$ {Centro de Astrobiolog\'ia, CSIC-INTA, Carretera de Torrej\'on a Ajalvir km-4, E-28850 Torrej\'on de Ardoz, Madrid, Spain}
  $^{12}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}
  $^{13}$ {Department of Physics, University of Oxford, Keble Road,
    Oxford OX13RH, UK} 
  $^{14}$ {Armagh Observatory, College Hill, Armagh BT619DG, UK}\\

  % \begin{acknowledgements}
  {   \tiny
    We thank P.~Crowther, J.~Heyl, M.\,C.\,Ramirez-Tannus, S.~N.~Shore,
    and S.\,Torres, and the HSTPROMO Collaboration
    for help and discussions. SdM acknowledges
    the European Unions Horizon 2020 research and innovation programme
    from the European Research Council (ERC), Grant agreement No. 715063
    and VHB the NRC-Canada Plaskett Fellowship. This work has made use of data from the European Space Agency (ESA) mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}).
  }
% \end{acknowledgements}

>>>>

\author[Renzo et al.]{M.~Renzo$^{1}$% \thanks{Corresponding author:
    % M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
  , S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2,3}$, I.~Platais$^{4}$,
  R.~P.~van~der~Marel$^{4,5}$,
  \newauthor{E.~Laplace$^{1}$, J.~M.~Bestenlehner$^{6}$, C.~J.~Evans$^{7}$,
    V.~H\'enault-Brunet$^{8}$,  S.~Justham$^{9,10,1}$,
  }
\newauthor{A.~de~Koter$^{1}$,
      N.~Langer$^{11}$, F. Najarro$^{12}$, H.~Sana$^{13}$, F.~R.~N.~Schneider$^{14}$, J.~S.~Vink$^{15}$}}

@@

\begin{keywords}
  stars: astrometry, kinematics and dynamics, individual: VFTS682
\end{keywords}
\vspace{-30pt}
\section{Introduction}
\label{sec:intro}

@@

 \hline     
      \multicolumn{5}{l}{\emph{HST proper motion of VFTS682 relative
      to R136}} \\
            \hline
      &$\delta\mu_\mathrm{RA, HST}$  &[$\masyr$] & $0.02\pm0.10$ & (5) \\
      &$\delta\mu_\mathrm{DEC, HST}$  &[$\masyr$] & $0.19\pm0.09$ &  (5) \\
       &$\delta\mu_\mathrm{HST}$  &[$\masyr$] & $0.19\pm0.09$ &  (5) \\
                  &$v_\mathrm{2D, HST}$  &[$\kms$] & $45\pm21$ & (5)\\  
                      &$\theta_\mathrm{HST}$  &[degrees] &   $-30_{-51}^{+24}$   & (1,6)\\  
                  %        &$v_\mathrm{2D, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
      \hline
      \multicolumn{5}{l}{\emph{Weighted average relative proper motion
      for VFTS682}}\\
      \hline
      &$\delta\mu_\mathrm{RA, avg}$  &[$\masyr$] & $0.08\pm0.07$ & (6) \\
      &$\delta\mu_\mathrm{DEC, avg}$  &[$\masyr$] & $0.14\pm0.07$ &  (6) \\
       &$\delta\mu_\mathrm{avg}$  &[$\masyr$] & $0.16\pm0.07$ &  (6) \\
      &$v_\mathrm{2D, avg}$  &[$\kms$] & $38\pm17$ & (6)\\
      % &$\theta_\mathrm{avg}$  &[degrees] &  \todo{TBD} $_{-}^{+}$   & (5)\\  
      \hline
       \multicolumn{5}{l}{\emph{Expected proper motion if ejected from
      R136 at age zero}} \\
      \hline
      &$v_\mathrm{2D}$  &[$\kms$] & $29\pm 6$ & (3) \\  
      &$\theta$  &[degrees] &  $\sim0$  & \\ 
       \hline
    \end{tabular}
    {\tiny The error on the RA and DEC positions, are of order
      $\sim$$0.01\masyr$ in \emph{Gaia}
      DR2. Assuming a distance of 50\,kpc, $1\masyr$ corresponds to $237\kms$.
$\rho(\mu_\mathrm{RA},\mu_\mathrm{DEC})$ is the
      correlation coefficient. The position angle
      $\theta$ is defined such that $\theta=0$ for radial motion
      away from R136. We neglect the error bars on
      $\langle\mu_\mathrm{RA}\rangle_\mathrm{R136}$ and $\langle\mu_\mathrm{DEC}\rangle_\mathrm{R136}$ to determine the uncertainty
      on $\theta_{Gaia}$.
      (1)~\cite{brown:18},
      (2)~\cite{henault-brunet:12},
      (3)~\cite{bestenlehner:11},
      (4)~\cite{lennon:18}, 
      (5)~\cite{platais:18} and
      (6)~{\color{blue}this study}.
    }
  \end{center}
  \label{tab:vfts682}
\end{table}

@@

\begin{figure}%[tbp]
  \centering
  \includegraphics[width=0.35\textwidth]{figures/dist_mu_region.pdf}\\
  \vspace*{-5pt}
  \includegraphics[width=0.35\textwidth]{figures/angle}
  % \caption{\emph{Top panel}: distribution of OB-type and Wolf-Rayet stars in proper
  %   motion relative to R136 from \emph{Gaia} DR2. VFTS682 is not an outlier, but
  %   its relative proper motion matches the expected value (green
  %   vertical band) if it were indeed
  %   ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. %The top axis shows the conversion to $\kms$
  %   % assuming a distance of 50\,kpc.
  %   \emph{Bottom panel}:  distribution of
  %   angles between the relative proper motion direction and the radial
  %   direction from the center of the cluster to the star. The error bars from \emph{Gaia} for VFTS682 are large, but
  %   the best value is in agreement with the hypothesis of dynamical
  %   ejection (green vertical band). The peak at $\theta\simeq\pi/2$ is due to stars
  %   belonging to NGC2060. In both
  %   panels, the dark blue histograms contain 317 
  %   stars with error bars smaller than $0.1\,\mathrm{mas \
  %     yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
  %   lighter blue histograms contain 36 stars with error bars smaller than $0.05\,\mathrm{mas \
  %   yr^{-1}}$. }
  \vspace{-5pt}
  \caption{Distribution of OB-type and Wolf-Rayet stars in proper
    motion relative to R136 (top panel) and proper motion position angle
    (bottom panel), from \emph{Gaia} DR2. Although VFTS682 is not an outlier, its relative
    proper motion matches the value expected for an early dynamical
    ejection \newtext{(see \Secref{sec:results})}. In both
    panels, the dark blue histograms contain 317 
    stars with error bars smaller than $0.1\,\mathrm{mas \
      yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc and the
    lighter blue histograms contain 36 stars with error bars smaller than $0.05\,\mathrm{mas \
    yr^{-1}}$. The peak at $\theta\simeq\pi/2$ in the bottom panel is due to stars
  belonging to NGC2060. }
\vspace{-10pt}
  \label{fig:dist}
\end{figure}

@@

\vspace{-30pt}
\section{The kinematics of VFTS682}
\label{sec:results}

@@

If confirmed, % it means that
isolated star formation is
\emph{not} required to explain the isolation of VFTS682. Its proper motion suggests that it was ejected from the cluster R136
$0.7\pm0.3$\,Myr ago, which is compatible with the evolutionary age of
the star. If the
cluster age \citep[$\lesssim2$\,Myr,][]{crowther:10, sabbi:12} is
indeed smaller than the shortest stellar lifetime
\citep[$\sim$3\,Myr,][]{brott:11,kohler:15, zapartas:17}, the ejection of VFTS682
from the disruption of a \newtext{massive} binary by a supernova is excluded. %  since the region is too young for stars
% to have experienced core-collapse already. 
The kinematic age we infer is smaller than the kinematic age of
$\sim$1.5\,Myr for VFTS16 found in \cite{lennon:18}, which indicates
that VFTS682 was ejected later than VFTS16, and potentially later than
VFTS72 too.

@@

\vspace*{-25pt}
\section*{Affiliations}
 \tiny
\noindent $^{1}${Astronomical Institute Anton Pannekoek, University of
    Amsterdam, 1098 XH Amsterdam, The Netherlands} 
  $^{2}$ {ESA, European Space Astronomy Centre, Apdo. de Correos 78,
    E-28691 Villanueva de la Ca\~nada, Madrid, Spain} 
 $^{3}$ {Instituto de Astrof\'isica de Canarias, E-38205 La Laguna, Tenerife, Spain}
 %Center for Astrophysical Sciences, 
$^4$ {Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}
  $^{5}$ {Space Telescope Science Institute, 3700 San Martin Drive,
    Baltimore, MD 21218, USA}
  $^{6}${Department of Physics and Astronomy, Hicks Building,
    Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}
  $^{7}${UK Astronomy Technology Centre, Royal Observatory Edinburgh, Blackford Hill, Edinburgh, EH9 3HJ, UK}
  $^{8}$ {National Research Council, Herzberg Astronomy \&
    Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7,
    Canada}
  $^{9}$ {School of Astronomy \& Space Science, University of the Chinese
    Academy of Sciences, Beijing 100012, China}
  $^{10}$ {National Astronomical Observatories, Chinese Academy of
    Sciences, Beijing 100012, China}
  $^{11}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn,
    Auf dem H\"ugel 71, 53121, Bonn, Germany}
  $^{12}$ {Centro de Astrobiolog\'ia, CSIC-INTA, Carretera de Torrej\'on a Ajalvir km-4, E-28850 Torrej\'on de Ardoz, Madrid, Spain}
  $^{13}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}
  $^{14}$ {Department of Physics, University of Oxford, Keble Road,
    Oxford OX13RH, UK} 
  $^{15}$ {Armagh Observatory, College Hill, Armagh BT619DG, UK}

  % \begin{acknowledgements}
  {   \tiny
    We thank P.~Crowther, J.~Heyl, M.\,C.\,Ramirez-Tannus, S.~N.~Shore,
    and S.\,Torres, and the HSTPROMO Collaboration
    for help and discussions. SdM acknowledges
    the European Unions Horizon 2020 research and innovation programme
    from the European Research Council (ERC), Grant agreement No. 715063
    and VHB the NRC-Canada Plaskett Fellowship. This study has used data from the ESA mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}).
  }
% \end{acknowledgements}
